---
title: "Final 2018"
author: "Maria Schweer-Collins"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nhanes 2015-2016 instructions

The code chunk below explains the nhanes and the other packages that are needed and how to load the data.

``` {r nhanes data, echo = TRUE}

install.packages("nhanesA")

library(nhanesA)
library(tidyverse)

```
The nhanesA package lets you import data directly from the CDC website on nhanes data. The code below will import data and I have created an object for it to be loaded to.
Below are examples of data from the 2015-2016 datasets.
You can also view the data but be prepared for your computer to take a while since it takes so long to download all the data.
It might be best to just view the data on the [nhanes website](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PAQ_I.htm) and then create an object with the specific variables we want to look at.

Now we don't have to load data through  daniel's method using the tidyverse `data_ex <- import(head("specific_folder", "dataset_name_ex" ))`

** Note: We could also look at visuals between the years since many of these variables reappear throughout all the years. **
[nhanes datasets by different years](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)

``` {r importing data, echo = TRUE}

# Example of body measures dataset from 2015-2016
bmx_i <- nhanes('BMX_I')

# Example of physical activity measures from 2015-2016
paq_i <- nhanes('PAQ_I')

View(bmx_i)
View(paq_i)

# bmx_i %>% 
  # select(put variables in here)
  
# Example of exporting data into csv files
write.csv(bmx_i, "bmx_i.csv")

```
