---
title             : "Final Project: Group 3"
shorttitle        : "Title"

author: 
  - name          : "Shaina Trevino"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Maria Schweer-Collins"
    affiliation   : "1"
  - name          : "Alejandra Garcia Isaza"
    affiliation   : "1"
  - name          : "Ernst-August Doelle"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "University of Oregon"
 

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Sign Up:
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)
library(rio)
library(here)
library(kableExtra) #If install is needed use - devtools::install_github("haozhu233/kableExtra")
theme_set(theme_minimal())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r data, echo = FALSE}

# Loading CSV files
duq_i <- import(here("data", "duq_i.csv"), 
                setclass = "tbl_df")
demo_i <- import(here("data", "demo_i.csv"),
                setclass = "tbl_df")
dpq_i <- import(here("data", "dpq_i.csv"),
                setclass = "tbl_df")
huq_i <- import(here("data", "huq_i.csv"),
                setclass = "tbl_df")
hiq_i <- import(here("data", "hiq_i.csv"),
                setclass = "tbl_df")
#merge files
d <- merge(duq_i,demo_i, by = "V1")
d1 <- merge(dpq_i, huq_i, by = "V1")
d2 <- merge(d, d1, by = "V1") 
fd <- merge(d2, hiq_i, by = "V1")

#select columns
td0 <- fd %>% 
  select(V1,DPQ010, DPQ020, DPQ030, DPQ040, DPQ050, DPQ060, DPQ070, DPQ080, DPQ090, DPQ100, RIDRETH3, RIAGENDR, DMDEDUC2, INDHHIN2, RIDAGEYR, DUQ260, DUQ300, DUQ340, DUQ430, DUQ210, HIQ011, HUQ090) 
#computing depression total score
td0 <- td0 %>%
  mutate(sum_depression = DPQ010 + DPQ020 + DPQ030 + DPQ040 + DPQ050 + DPQ060 + DPQ070 + DPQ080 + DPQ090)
#replace value with NA
#install.packages("naniar")
 td <- td0 %>% 
   naniar::replace_with_na(replace = list(DUQ340 = 999)) %>% 
   mutate(age_stim = (DUQ260 + DUQ340) / 2) 
# Labels for categorical variables
td <- td %>% 
  mutate(ethnicity = case_when(RIDRETH3 == 1 ~ "Hispanic", #Shaina changed these for paper
                               RIDRETH3 == 2 ~ "Hispanic", 
                               RIDRETH3 == 3 ~ "White",
                               RIDRETH3 == 4 ~ "Black",
                               RIDRETH3 == 6 ~ "Asian",
                               RIDRETH3 == 7 ~ "Other Race - Including Multi-Racial",
                               TRUE ~ NA_character_)) 
td <- td %>% 
  mutate(sex = case_when(RIAGENDR == 1 ~ "Male", 
                         RIAGENDR == 2 ~ "Female", 
                         TRUE ~ NA_character_)) 
td <- td %>% 
  mutate(education = case_when(DMDEDUC2 == 1 ~ "Less than 9th grade",
                               DMDEDUC2 == 2 ~ "9-12th grade (no diploma)",
                               DMDEDUC2 == 3 ~ "High school graduate/GED or equivalent",
                               DMDEDUC2 == 4 ~ "Some college of AA degree",
                               DMDEDUC2 == 5 ~ "College graduate of above", 
                               DMDEDUC2 == 7 ~ "Refused",
                               DMDEDUC2 == 9 ~ "Don't know",
                               TRUE ~ NA_character_))
td <- td %>% 
  mutate(income = case_when(INDHHIN2 == 1 ~ "$0 - $4,999",
                            INDHHIN2 == 2 ~ "$5,000 to $9,999",
                            INDHHIN2 == 3 ~ "10,000 to $14,999",
                            INDHHIN2 == 4 ~ "$15,000 to $19,999",
                            INDHHIN2 == 5 ~ "$20,000 to $24,999",
                            INDHHIN2 == 6 ~ "$25,000 to $34,999",
                            INDHHIN2 == 7 ~ "$35,000 to $44,999",
                            INDHHIN2 == 8 ~ "$45,000 to $54,999",
                            INDHHIN2 == 9 ~ "$55,000 to $64,999",
                            INDHHIN2 == 10 ~ "$65,000 to $74,999",
                            INDHHIN2 == 12 ~ NA_character_,
                            INDHHIN2 == 13 ~ NA_character_,
                            INDHHIN2 == 14 ~ "$75,000 to $99,999",
                            INDHHIN2 == 15 ~ "$100,000 and over",
                            INDHHIN2 == 77 ~ "Refused",
                            INDHHIN2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))
#RIDAGEYR is half categorical (0-79 are values and 80 is 80 and older)
td <- td %>% 
  mutate(insurance = case_when(HIQ011 == 1 ~ "Yes",
                               HIQ011 == 2 ~ "No",
                               HIQ011 == 7 ~ "Refused",
                               HIQ011 == 9 ~ "Don't know",
                               TRUE ~ NA_character_))
td <- td %>% 
  mutate(mental_health = case_when(HUQ090 == 1 ~ "Yes",
                                   HUQ090 == 2 ~ "No",
                                   HUQ090 == 7 ~ "Refused",
                                   HUQ090 == 9 ~ "Don't know",
                                   TRUE ~ NA_character_))
td <- td %>% 
  mutate(rehab = case_when(DUQ430 == 1 ~ "Yes",
                           DUQ430 == 2 ~ "No",
                           DUQ430 == 7 ~ "Refused",
                           DUQ430 == 9 ~ "Don't know",
                           TRUE ~ NA_character_))
#select and rename
td1 <- td %>% 
  select(-2:-15, -HIQ011, -HUQ090, -DUQ430, income) %>% 
  rename(age_marijuana = DUQ210, 
         age_cocaine = DUQ260, 
         age_heroin = DUQ300, 
         age_meth = DUQ340)
#reorder
td1 <- td1 %>% 
  select(1:6, age_stim, sum_depression, rehab, mental_health, insurance, income, 7:10) %>%
  rename(age = RIDAGEYR,
         id = V1)
```

#Introduction
Alejandra will do depression and health disparities (short 1 parapraph)
Maria will do drug use onset and health disparities (short 1 paragraph)

# Methods
Sign up:

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants
Sign up:

##Measures
sign up:
Alejandra will describe the measures for depression, use of mental health services, and insurance coverage. 

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
sign up:

# Results
We should use inline code here
sign up: 

Alejandra will do the description of her plot (depression, health insurance coverage, visit to mental health, broken down by ethnicities). 

```{r plot1, echo = FALSE, message = FALSE}

td1 %>%
  filter(mental_health != "Don't know") %>% 
  filter(insurance == "Yes" | insurance == "No") %>% 
  mutate(ethnicity = fct_infreq(ethnicity),
         mental_health = fct_infreq(mental_health)) %>%
  group_by(ethnicity, mental_health) %>%
  ggplot(aes(x = ethnicity, y = sum_depression)) +
  geom_col(aes(fill = mental_health),
           position = "dodge") +
   labs(title = "Total Depression Score by Ethnicity and Usage of Mental Health Services",
       subtitle = "Faceted by Insurance Coverage",
         caption = "Source: NHANES",
         x = "Ethnicity",
         y = "Total Depression Score",
       fill = "Usage of Mental Health Services") +
  facet_wrap(~insurance) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



Shaina will describe her plot of drug use by ethnicities
```{r ST_plot, echo = FALSE, fig.height = 4, dev = 'svg'}
pd_st <- td1 %>% 
           select(id, age_stim, age_heroin, sum_depression, rehab, ethnicity) %>%
           rename(age_Stimulants = age_stim,
                  age_Opioids = age_heroin) %>% 
           gather(drug, age_use, age_Stimulants:age_Opioids) %>% 
           separate(drug, c("dis", "drug"), sep = "_") %>% 
           select(-dis) %>% 
           filter(ethnicity != "Other Race - Including Multi-Racial") %>% 
           na.omit(age_use)

ggplot(pd_st, aes(x = fct_infreq(ethnicity), y = age_use)) +
  geom_violin(aes(fill = ethnicity), alpha = .8) +
  facet_wrap (~ drug) +
  labs(title = "Age of First Use by Ethnicity", 
       subtitle = "Based on two drug types (opioids and stimulants)",
       caption = "Source: NHANES",
       x = "Ethnicity",
       y = "Age at first use") #get rid of legend
```


```{r ST_plot1, echo = FALSE, fig.height = 4, dev = 'svg'}
td1 %>% 
  select(1, 3:6, sum_depression, ethnicity) %>% 
  gather(drug, age_use, 2:4) %>% 
  separate(drug, c("dis", "drug"), sep = "_") %>% 
  select(-dis) %>% 
  na.omit() %>% 
  ggplot(aes(x = age_use, y = sum_depression, color = ethnicity)) +
  geom_point(size = .75) +
  geom_smooth(method = "lm", se = FALSE, size = 1) +
  facet_wrap(~drug)
  labs(title = "Association between Age at First Use and Depression", 
       subtitle = "Colored to show differences among drug types",
       caption = "Source: NHANES",
       x = "Age at first use ",
       y = "Total depression score") 

```
# Discussion
sign up:

Insert one data visualization -- we are using two, I believe Alejandra's viz. and one from Shaina
Insert Table -- JP?


\newpage

# References
sign up:
Alejandra will include the references used in the intro. 
```{r create_r-references}
r_refs(file = "r-references.bib")

  
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
