---
title: "Final 2018"
author: "Maria Schweer-Collins"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nhanes 2015-2016 instructions

The code chunk below explains the nhanes and the other packages that are needed and how to load the data.

``` {r nhanes data, echo = FALSE}

# install.packages("nhanesA")

library(nhanesA)
library(tidyverse)
library(rio)
library(here)
library(janitor)
theme_set(theme_classic())
```
The nhanesA package lets you import data directly from the CDC website on nhanes data. The code below will import data and I have created an object for it to be loaded to.

Below are examples of data from the 2015-2016 datasets.

You can also view the data but be prepared for your computer to take a while since it takes so long to download all the data.

It might be best to just view the data on the [nhanes website](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PAQ_I.htm) and then create an object with the specific variables we want to look at.

** Note: We could also look at visuals between the years since many of these variables reappear throughout all the years. **

[nhanes datasets by different years](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)

``` {r load_data, echo = TRUE, eval = FALSE}

# drug use dataset from 2015-2016
duq_i <- nhanes('DUQ_I')

# Demographics measures from 2015-2016
demo_i <- nhanes('DEMO_I')

# Depression measure
dpq_i <- nhanes('DPQ_I')

# Mental health services measure
huq_i <- nhanes('HUQ_I')

# Health insurance measure
hiq_i <- nhanes('HIQ_I')


# Example of exporting data into csv files
write.csv(duq_i, "duq_i.csv")
write.csv(demo_i, "demo_i.csv")
write.csv(dpq_i, "dpq_i.csv")
write.csv(huq_i, "huq_i.csv")
write.csv(hiq_i, "hiq_i.csv")

```

```{r import_data}
# Loading CSV files
duq_i <- import(here("data", "duq_i.csv"), 
                setclass = "tbl_df")
demo_i <- import(here("data", "demo_i.csv"),
                setclass = "tbl_df")
dpq_i <- import(here("data", "dpq_i.csv"),
                setclass = "tbl_df")
huq_i <- import(here("data", "huq_i.csv"),
                setclass = "tbl_df")
hiq_i <- import(here("data", "hiq_i.csv"),
                setclass = "tbl_df")

#merge files
d <- merge(duq_i,demo_i, by = "V1")
d1 <- merge(dpq_i, huq_i, by = "V1")
d2 <- merge(d, d1, by = "V1") 
fd <- merge(d2, hiq_i, by = "V1") # fd stands for final data

View(fd)
            
# Clean variables - DOESNT WORK YETTTTT!!!!!!
#fd%>% 
  #clean_names()
```

```{r tidy}

# Select columns (variables)
td <-  fd %>% 
  select(V1,DPQ010, DPQ020, DPQ030, DPQ040, DPQ050, DPQ060, DPQ070, DPQ080, DPQ090, DPQ100, RIDRETH3, RIAGENDR, DMDEDUC2, INDHHIN2, RIDAGEYR, DUQ260, DUQ300, DUQ340, DUQ430, DUQ210, HIQ011, HUQ090)

  
# Evaluation of missing data

td_without_missing <- na.omit(td) # if missing data is removed, we would just have a sample of 17 people. When doing calculation, we will use the na.rm = TRUE. 

View(td_without_missing) 

# The DUQ variables are mostly missing data, I don't know if it makes sense to use variables. 

# Computing depression total score

td <-td %>%
  mutate(sum_depression = DPQ010+ DPQ020+ DPQ030+DPQ040+ DPQ050+ DPQ060+ DPQ070+ DPQ080 + DPQ090)
View(td) #DA changed this because the original code messed up the data frame

#td %>% 
  #mutate(sum_depress = rowSums(.[ ,15:20], na.rm = TRUE)) #can use rowmeans

# Computing stimulant composite and converting demographic variables to factors
 td <- td %>% 
   mutate(age_stim = (DUQ260 + DUQ340) / 2) %>% #should be using group by and mean but cant yet because data is not tidy, Per DA suggestion, we need to tidy so that we have column for item and column for domain (e.g., depression) then you can group by and summarize means. 
   mutate(RIDRETH3 = factor(RIDRETH3)) %>%
   mutate(RIAGENDR = factor(RIAGENDR))

# Labels for categorical variables
td <- td %>% 
  mutate(ethnicity = case_when(RIDRETH3 == 1 ~ "Mexican American", 
                 RIDRETH3 == 2 ~ "Other Hispanic", 
                 RIDRETH3 == 3 ~ "Non-Hispanic White",
                 RIDRETH3 == 4 ~ "Non-Hispanic Black",
                 RIDRETH3 == 6 ~ "Non-Hispanic Asian",
                 RIDRETH3 == 7 ~ "Other Race - Including Multi-Racial",
                 TRUE ~ NA_character_)) 

td <- td %>% 
  mutate(sex = case_when(RIAGENDR == 1 ~ "Male", 
                 RIAGENDR == 2 ~ "Female", 
                 TRUE ~ NA_character_)) 

td <- td %>% 
  mutate(education = case_when(DMDEDUC2 == 1 ~ "Less than 9th grade",
                               DMDEDUC2 == 2 ~ "9-12th grade (no diploma)",
                               DMDEDUC2 == 3 ~ "High school graduate/GED or equivalent",
                               DMDEDUC2 == 4 ~ "Some college of AA degree",
                               DMDEDUC2 == 5 ~ "College graduate of above", 
                               DMDEDUC2 == 7 ~ "Refused",
                               DMDEDUC2 == 9 ~ "Don't know",
                               TRUE ~ NA_character_))

td <- td %>% 
  mutate(income = case_when(INDHHIN2 == 1 ~ "$0 - $4,999",
                            INDHHIN2 == 2 ~ "$5,000 to $9,999",
                            INDHHIN2 == 3 ~ "10,000 to $14,999",
                            INDHHIN2 == 4 ~ "$15,000 to $19,999",
                            INDHHIN2 == 5 ~ "$20,000 to $24,999",
                            INDHHIN2 == 6 ~ "$25,000 to $34,999",
                            INDHHIN2 == 7 ~ "$35,000 to $44,999",
                            INDHHIN2 == 8 ~ "$45,000 to $54,999",
                            INDHHIN2 == 9 ~ "$55,000 to $64,999",
                            INDHHIN2 == 10 ~ "$65,000 to $74,999",
                            INDHHIN2 == 12 ~ "$20,000 and over",## duplicates/composites from above
                            INDHHIN2 == 13 ~ "Under $20,000",## duplicates/composites from above
                            INDHHIN2 == 14 ~ "$75,000 to $99,999",
                            INDHHIN2 == 15 ~ "$100,000 and over",
                            INDHHIN2 == 77 ~ "Refused",
                            INDHHIN2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))
#RIDAGEYR is half categorical (0-79 are values and 80 is 80 and older)



 
```

```{r exploratory_visuals}
#Maria's visualization 

#prepare data

gender <- as.factor(td$RIAGENDR)

#Create boxplot of total depression scores by gender

plot1 <- ggplot(td, aes(gender, sum_depression))
plot1 + geom_boxplot(varwidth=T, fill="plum") + 
    labs(title="Box plot", 
         subtitle="Total Depression Score by Gender",
         caption="Source: NHANES",
         x="Gender",
         y="Total Depression Score")

```








