---
title: "Final 2018"
author: "Maria Schweer-Collins"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nhanes 2015-2016 instructions

The code chunk below explains the nhanes and the other packages that are needed and how to load the data.

``` {r nhanes data, echo = FALSE}

# install.packages("nhanesA")

library(nhanesA)
library(tidyverse)
library(rio)
library(here)
library(janitor)

```
The nhanesA package lets you import data directly from the CDC website on nhanes data. The code below will import data and I have created an object for it to be loaded to.
Below are examples of data from the 2015-2016 datasets.
You can also view the data but be prepared for your computer to take a while since it takes so long to download all the data.
It might be best to just view the data on the [nhanes website](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PAQ_I.htm) and then create an object with the specific variables we want to look at.

Now we don't have to load data through  daniel's method using the tidyverse `data_ex <- import(head("specific_folder", "dataset_name_ex" ))`

** Note: We could also look at visuals between the years since many of these variables reappear throughout all the years. **
[nhanes datasets by different years](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)

``` {r load_data, echo = TRUE, eval = FALSE}

# drug use dataset from 2015-2016
duq_i <- nhanes('DUQ_I')

# Demographics measures from 2015-2016
demo_i <- nhanes('DEMO_I')

# Depression measure
dpq_i <- nhanes('DPQ_I')

# Mental health services measure
huq_i <- nhanes('HUQ_I')

# Health insurance measure
hiq_i <- nhanes('HIQ_I')


# Example of exporting data into csv files
write.csv(duq_i, "duq_i.csv")
write.csv(demo_i, "demo_i.csv")
write.csv(dpq_i, "dpq_i.csv")
write.csv(huq_i, "huq_i.csv")
write.csv(hiq_i, "hiq_i.csv")

```

```{r import_data}
# Loading CSV files
duq_i <- import(here("data", "duq_i.csv"), 
                setclass = "tbl_df")
demo_i <- import(here("data", "demo_i.csv"),
                setclass = "tbl_df")
dpq_i <- import(here("data", "dpq_i.csv"),
                setclass = "tbl_df")
huq_i <- import(here("data", "huq_i.csv"),
                setclass = "tbl_df")
hiq_i <- import(here("data", "hiq_i.csv"),
                setclass = "tbl_df")

#merge files
d <- merge(duq_i,demo_i, by = "V1")
d1 <- merge(dpq_i, huq_i, by = "V1")
d2 <- merge(d, d1, by = "V1") 
fd <- merge(d2, hiq_i, by = "V1")
            
# Clean variables - DOESNT WORK YETTTTT!!!!!!
#fd%>% 
  #clean_names()

# Select columns (variables)
duq_i %>% 
  

```
