---
title: "Final 2018"
author: "Maria Schweer-Collins"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nhanes 2015-2016 instructions

The code chunk below explains the nhanes and the other packages that are needed and how to load the data.

``` {r nhanes data, echo = FALSE}

# install.packages("nhanesA")
# install.packages("naniar")

library(nhanesA)
library(tidyverse)
library(rio)
library(here)
theme_set(theme_classic())
```
The `nhanesA` package lets you import data directly from the CDC website on nhanes data. The code below will import data and I have created an object for it to be loaded to.

Below are examples of data from the 2015-2016 datasets.

You can also view the data but be prepared for your computer to take a while since it takes so long to download all the data.

It might be best to just view the data on the [nhanes website](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PAQ_I.htm) and then create an object with the specific variables we want to look at.

** Note: We could also look at visuals between the years since many of these variables reappear throughout all the years. **

[nhanes datasets by different years](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)

``` {r load_data, echo = TRUE, eval = FALSE}

# Don't need to run this - it is only prep work
#
# drug use dataset from 2015-2016

duq_i <- nhanes('DUQ_I')

# Demographics measures from 2015-2016
demo_i <- nhanes('DEMO_I')

# Depression measure
dpq_i <- nhanes('DPQ_I')

# Mental health services measure
huq_i <- nhanes('HUQ_I')

# Health insurance measure
hiq_i <- nhanes('HIQ_I')


# Example of exporting data into csv files
write.csv(duq_i, "duq_i.csv")
write.csv(demo_i, "demo_i.csv")
write.csv(dpq_i, "dpq_i.csv")
write.csv(huq_i, "huq_i.csv")
write.csv(hiq_i, "hiq_i.csv")

```

```{r import_data}

# Loading CSV files
duq_i <- import(here("data", "duq_i.csv"), 
                setclass = "tbl_df")
demo_i <- import(here("data", "demo_i.csv"),
                setclass = "tbl_df")
dpq_i <- import(here("data", "dpq_i.csv"),
                setclass = "tbl_df")
huq_i <- import(here("data", "huq_i.csv"),
                setclass = "tbl_df")
hiq_i <- import(here("data", "hiq_i.csv"),
                setclass = "tbl_df")

#merge files
d <- merge(duq_i,demo_i, by = "V1")
d1 <- merge(dpq_i, huq_i, by = "V1")
d2 <- merge(d, d1, by = "V1") 
fd <- merge(d2, hiq_i, by = "V1") # fd stands for final merged data

            
```

```{r tidy}

# Select columns (variables)
td <-  fd %>% 
  select(V1,DPQ010, DPQ020, DPQ030, DPQ040, DPQ050, DPQ060, DPQ070, DPQ080, DPQ090, DPQ100, RIDRETH3, RIAGENDR, DMDEDUC2, INDHHIN2, RIDAGEYR, DUQ260, DUQ300, DUQ340, DUQ430, DUQ210, HIQ011, HUQ090)

  
# Evaluation of missing data

td_without_missing <- na.omit(td) # if all missing data is removed, we would just have a sample of 17 people. When doing calculation, we will use the na.rm = TRUE. 

View(td_without_missing) 

```

```{r composite}

# Computing depression total score

td <-td %>%
  mutate(sum_depression = DPQ010 + DPQ020 + DPQ030 + DPQ040 + DPQ050 + DPQ060 + DPQ070 + DPQ080 + DPQ090) #DPQ100 is not included because the response scale changes. 


#td %>% 
  #mutate(sum_depress = rowSums(.[ ,15:20], na.rm = TRUE)) # This is an alternative way to do it that DA suggested. Not used here, though. Can also use rowmeans in this line of code. Keeping for future reference.

# Computing stimulant composite and converting demographic variables to factors
# Install package first, if needed (uncomment and run code below):

#install.packages("naniar")
 td <- td %>% 
   naniar::replace_with_na(replace = list(DUQ340 = 999)) %>% 
   mutate(age_stim = (DUQ260 + DUQ340) / 2) 

```

```{r factors}

td <- td %>% #Not sure if we still need this if we are creating new variables that are already characters?
  mutate(RIDRETH3 = factor(RIDRETH3)) %>%
  mutate(RIAGENDR = factor(RIAGENDR))

# Labels for categorical variables
td <- td %>% 
  mutate(ethnicity = case_when(RIDRETH3 == 1 ~ "Mexican American", 
                 RIDRETH3 == 2 ~ "Other Hispanic", 
                 RIDRETH3 == 3 ~ "Non-Hispanic White",
                 RIDRETH3 == 4 ~ "Non-Hispanic Black",
                 RIDRETH3 == 6 ~ "Non-Hispanic Asian",
                 RIDRETH3 == 7 ~ "Other Race - Including Multi-Racial",
                 TRUE ~ NA_character_)) 

td <- td %>% 
  mutate(sex = case_when(RIAGENDR == 1 ~ "Male", 
                 RIAGENDR == 2 ~ "Female", 
                 TRUE ~ NA_character_)) 

td <- td %>% 
  mutate(education = case_when(DMDEDUC2 == 1 ~ "Less than 9th grade",
                               DMDEDUC2 == 2 ~ "9-12th grade (no diploma)",
                               DMDEDUC2 == 3 ~ "High school graduate/GED or equivalent",
                               DMDEDUC2 == 4 ~ "Some college of AA degree",
                               DMDEDUC2 == 5 ~ "College graduate of above", 
                               DMDEDUC2 == 7 ~ "Refused",
                               DMDEDUC2 == 9 ~ "Don't know",
                               TRUE ~ NA_character_))

# JP - adjusted the code to simply remove the answer choices that did not make sense
td <- td %>% 
  mutate(income = case_when(INDHHIN2 == 1 ~ "$0 - $4,999",
                            INDHHIN2 == 2 ~ "$5,000 to $9,999",
                            INDHHIN2 == 3 ~ "10,000 to $14,999",
                            INDHHIN2 == 4 ~ "$15,000 to $19,999",
                            INDHHIN2 == 5 ~ "$20,000 to $24,999",
                            INDHHIN2 == 6 ~ "$25,000 to $34,999",
                            INDHHIN2 == 7 ~ "$35,000 to $44,999",
                            INDHHIN2 == 8 ~ "$45,000 to $54,999",
                            INDHHIN2 == 9 ~ "$55,000 to $64,999",
                            INDHHIN2 == 10 ~ "$65,000 to $74,999",
                            INDHHIN2 == 12 ~ NA_character_,
                            INDHHIN2 == 13 ~ NA_character_,
                            INDHHIN2 == 14 ~ "$75,000 to $99,999",
                            INDHHIN2 == 15 ~ "$100,000 and over",
                            INDHHIN2 == 77 ~ "Refused",
                            INDHHIN2 == 99 ~ "Don't know",
                            TRUE ~ NA_character_))
#RIDAGEYR is half categorical (0-79 are values and 80 is 80 and older)

td <- td %>% 
  mutate(insurance = case_when(HIQ011 == 1 ~ "Yes",
                               HIQ011 == 2 ~ "No",
                               HIQ011 == 7 ~ "Refused",
                               HIQ011 == 9 ~ "Don't know",
                               TRUE ~ NA_character_))

td <- td %>% 
  mutate(mental_health = case_when(HUQ090 == 1 ~ "Yes",
                            HUQ090 == 2 ~ "No",
                            HUQ090 == 7 ~ "Refused",
                            HUQ090 == 9 ~ "Don't know",
                            TRUE ~ NA_character_))

td <- td %>% 
  mutate(rehab = case_when(DUQ430 == 1 ~ "Yes",
                           DUQ430 == 2 ~ "No",
                           DUQ430 == 7 ~ "Refused",
                           DUQ430 == 9 ~ "Don't know",
                           TRUE ~ NA_character_))

```

```{r select_rename}
td1 <- td %>% 
  select(-2:-15, -HIQ011, -HUQ090, -DUQ430, income) %>% 
  rename(age_marijuana = DUQ210, age_cocaine = DUQ260, age_heroin = DUQ300, age_meth = DUQ340)

#reorder
td1 <- td1 %>% 
  select(1:6, age_stim, sum_depression, rehab, mental_health, insurance, income, 7:10) %>%
  rename(age = RIDAGEYR,
         id = V1) 
View(td1)
```

``` {r age id and income}
# JP's work
# Im not going to change age or recode it since I think it acts as another integer and does not affect the findings. It just shows it as 80+.
# Maria moved age and id rename up to the tidying above


td1 %>%  
  select(age) %>% 
  ggplot(aes(age)) +
  geom_histogram()

td1 %>%
  count(income)

```

```{r exploratory_visuals}
#Maria's visualization 

#prepare data

gender <- as.factor(td$RIAGENDR)

#Create boxplot of total depression scores by gender
# Maria updated to include the new td1

plot1 <- ggplot(td1, aes(sex, sum_depression))
plot1 + geom_boxplot(varwidth = T, fill = "plum") + 
    labs(title = "Box plot", 
         subtitle = "Total Depression Score by Gender",
         caption = "Source: NHANES",
         x = "Sex",
         y = "Total Depression Score") 

```

```{r ST_plot}
td1 %>% 
  select(1, 3:6, rehab) %>% 
  gather(drug, age, 2:4) %>% 
  separate(drug, c("dis", "drug"), sep = "_") %>% 
  select(-dis) %>% 
  na.omit() %>% 

ggplot(aes(x = age)) +
  geom_histogram(aes(fill = rehab)) +
  facet_wrap (~ drug)


```

```{r ST_plot1}
td1 %>% 
  select(1, 3:6, sum_depression) %>% 
  gather(drug, age, 2:4) %>% 
  separate(drug, c("dis", "drug"), sep = "_") %>% 
  select(-dis) %>% 
  na.omit() %>% 
  
  ggplot(aes(x = sum_depression, y = age, color = drug)) +
  geom_point(size = .75) +
  geom_smooth(method = "lm", se = FALSE, size = 1.25)
```
```{r plot by ethnicity}

td1 %>% 
  mutate(ethnicity = fct_infreq(ethnicity)) %>%
  group_by(ethnicity) %>%
  ggplot(aes(x = ethnicity, y = sum_depression)) +
  geom_col(fill = "cornflowerblue") +
  coord_flip()

```


```{r descriptives}
# age of first drug use by ethnicity plot 
# We will also visualize reported usage of mental health services in the past year and health insurance coverage by racial/ethnic group


# Reviewers, please let us know if you know how to get rid of "Refused" and
# "Don't know" from our plots

td1 %>%
  mutate(ethnicity = fct_infreq(ethnicity),
         mental_health = fct_infreq(mental_health)) %>%
    group_by(ethnicity, mental_health) %>%
  ggplot(aes(x = ethnicity, y = sum_depression)) +
  geom_col(aes(fill = mental_health),
           position = "dodge") +
  facet_wrap(~insurance) +
  coord_flip()

```

``` {r summarizing depresssion and age}
# summarizing data for depression and ages of drug uses across the different ethnicities
# tables included
# also have whether or not they used mental health services.

sum_td1 <- td1 %>% 
  group_by(ethnicity) %>% 
  summarize(depress_avg = mean(sum_depression, na.rm = TRUE),
            depress_sd = sd(sum_depression, na.rm = TRUE),
            cocaine_avg = mean(age_cocaine, na.rm = TRUE),
            cocaine_sd = sd(age_cocaine, na.rm = TRUE),
            heroin_avg = mean(age_heroin, na.rm = TRUE),
            heroin_sd = sd(age_heroin, na.rm = TRUE),
            meth_avg = mean(age_meth, na.rm = TRUE),
            meth_sd = mean(age_meth, na.rm = TRUE),
            marijuana_avg = mean(age_marijuana, na.rm = TRUE),
            marijuana_sd = sd(age_marijuana, na.rm = TRUE))

knitr::kable(sum_td1, format = "html", digits = round(2))

sum_td1_mental_no <- td1 %>% 
  group_by(ethnicity) %>%
  filter(mental_health == "No") %>% 
  summarize(depress_avg = mean(sum_depression, na.rm = TRUE),
            depress_sd = sd(sum_depression, na.rm = TRUE),
            cocaine_avg = mean(age_cocaine, na.rm = TRUE),
            cocaine_sd = sd(age_cocaine, na.rm = TRUE),
            heroin_avg = mean(age_heroin, na.rm = TRUE),
            heroin_sd = sd(age_heroin, na.rm = TRUE),
            meth_avg = mean(age_meth, na.rm = TRUE),
            meth_sd = mean(age_meth, na.rm = TRUE),
            marijuana_avg = mean(age_marijuana, na.rm = TRUE),
            marijuana_sd = sd(age_marijuana, na.rm = TRUE))

knitr::kable(sum_td1_mental_no, format = "html", digits = round(2))

sum_td1_mental_yes <- td1 %>% 
  group_by(ethnicity) %>%
  filter(mental_health == "Yes") %>% 
  summarize(depress_avg = mean(sum_depression, na.rm = TRUE),
            depress_sd = sd(sum_depression, na.rm = TRUE),
            cocaine_avg = mean(age_cocaine, na.rm = TRUE),
            cocaine_sd = sd(age_cocaine, na.rm = TRUE),
            heroin_avg = mean(age_heroin, na.rm = TRUE),
            heroin_sd = sd(age_heroin, na.rm = TRUE),
            meth_avg = mean(age_meth, na.rm = TRUE),
            meth_sd = mean(age_meth, na.rm = TRUE),
            marijuana_avg = mean(age_marijuana, na.rm = TRUE),
            marijuana_sd = sd(age_marijuana, na.rm = TRUE))

knitr::kable(sum_td1_mental_yes, format = "html", digits = round(2))

```

```{r exploratory spreading}
# demographics_ex <- td1 %>%
#   group_by(sex, income, ethnicity, age) %>% 
#   summarize(age_avg = mean(age),
#             age_sd = sd(age),
#             n = n())
```


